### 2022-01-27 aistats-submission/rebuttal-aistats/discussion.tex

% !TEX root = aistats2022_author_response.tex

\SLJ{%
Yes, kudos, all the reviewers seem to have found the paper interesting and have appreciated the writing! :) The only reason for the 4 and 5 is that they think that it is not a "completed result" and are unsure whether this is suitable.
~\\~\\
As an AC in machine learning, I disagree, and indeed there is no such notion about this on the call for papers. And importantly, there is the emphasis on "promoting the exchange of ideas" about ML and statistics in the summary for call for papers, nowhere does it say that these ideas should be "complete results"...
~\\~\\
You could also mention in the rebuttal is that each time you talked to people in optimization or statistics, they were interested in knowing the answers to these questions, and so this is why we decided to publish as is, to stimulate interest in these directions and get the help of the community to solve this hard problem...
}